---
title: "Elevation Data Cleaning"
author: "Adam Sparks"
date: "Oct 9, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Overview

## Specific Goal
Make clean data files for all studies for our 1st Elevation publication. 

## Project ID
Study labels are different in various places.

Study # in paper: Sparks files, OSF, Monkey/Gizmo & name

Study 1: 01a1 Free Response, NA, Monkey

Study 2: 01a2 Free Response, , Monkey

Study 3: 01b1 Attitude Emotion Behavior,  , Monkey

Study 4: Simple Elevation 1, , Gizmo

Study 5: Simple Elevation 2, , Gizmo

Study 6: Simple Elevation 3, , Gizmo

Study 7: Simple Elevation 4, , Gizmo

In Sparks personal files, 4, 5, 6, 7 currently "Simple Elevation 1, 2, 3, 4". On OSF, I think they're 1, 2, 4, 5 (with an incomplete field study as #3)

#Data Cleaning
## Data preparation in spreadsheet software

Notes about this step included here.

###Study1
- SurveyMonkey
- Delete unused columns (F:I)
- Create short variable names
- political: 1-ext. lib., 7-ext. consv.

###Study2
- SurveyMonkey
- check: pro2, anti3
- Delete columns F:I
- Create short variable names
- political: 1-ext. lib., 7-ext. consv.

###Study3
- SurveyMonkey
- Short names differ from other studies because there are so many (could come back and reconcile the names)
- did some attention-check-based cleaning in excel (could redo in R)

###Study 4
- Delete unused columns
- Create short variable names
- PartID 443 device type -- correction
- Delete Ps with failed comprehension checks (n=6)
- Word count with =IF(ISBLANK(AV10),0,LEN(TRIM(AV10))-LEN(SUBSTITUTE(AV10," ",""))+1)

###Study 5
- Delete unused columns
- Create short variable names
- Word count with =IF(ISBLANK(AV10),0,LEN(TRIM(AV10))-LEN(SUBSTITUTE(AV10," ",""))+1)

###Study 6
- delete unused columns
- paste (transposed) column headers to new tab
- create short variable names
    - "Right now, I feel that" attitude items cannot be distinguished based on the standard results output because the order of variables on the saved PDF is randomized. I solved this by looking at individual level results, which report the full text of the multiple choice selection for these items.

### Study 7
- delete columns D:M (unused + unique MTurk links that could be identifiers)
- paste (transposed) column headers to new tab & create short variable names: "study7_variablenames.csv"

## Load data
Here I load the .csv raw files.
```{r}
study1 = read.csv("study1_rawdata.csv", header = TRUE)
study2 = read.csv("study2_rawdata.csv", header = TRUE)
study3 = read.csv("study3_rawdata.csv", header = TRUE)
study4 = read.csv("study4_rawdata.csv", header = TRUE)
study5 = read.csv("study5_rawdata.csv", header = TRUE)
study6 = read.csv("study6_rawdata.csv", header = TRUE)
study7 = read.csv("study7_rawdata.csv", header = TRUE)

#over life of this project I've changed how I deal with variable names. I plan to import separate files of short/long names going forward.

variables_study1 = read.csv("study1_variablenames.csv", header = TRUE)
attributes(study1)$variable.labels = variables_study1[, 2:3]  #i thought those would show as subheaders in the data frame viewer

variables_study2 = read.csv("study2_variablenames.csv", header = TRUE)
variables_study3 = read.csv("study3_variablenames.csv", header = FALSE)
variables_study7 = read.csv("study7_variablenames.csv", header = FALSE)
names(study7) = variables_study7[, 1] #use short names; column 2 has the full items

#these studies include ratings collected separately
study4 = merge (study4, read.csv("study4_helpfulness.csv"), by = "Response.ID")
study5 = merge (study5, read.csv("study5_helpfulness.csv"), by = "Response.ID")
study6 = merge (study6, read.csv("study6_helpfulness.csv"), by = "Response.ID")
study7 = merge (study7, read.csv("study7_helpfulness.csv"), by = "Response.ID")

# starting sample size
# did I do any excel exclusions that wouldn't be noted here? (probably study 3)
starting_sample = list(
  "Study 1" = dim(study1)[1],
  "Study 2" = dim(study2)[1],
  "Study 3" = dim(study3)[1],
  "Study 4" = dim(study4)[1],
  "Study 5" = dim(study5)[1],
  "Study 6" = dim(study6)[1],
  "Study 7" = dim(study7)[1]
  )
```

## Variables of Interest
For Study 3 there are so many items that (at least for now) I have just numbered them rather than givng them a handy short label.
The variables for studies 4 and 5 were nearly identical, with only a few DVs differing. 
Studies 6 & 7 have many more items because of the repeated measures aspect of the designs.

### Identifying conditions
#### Study 1
Study 1 included three video conditions: 1-pro, 2-anti, 3-neut.
```{r}
study1$condition = substring(study1$condition, nchar(as.character(study1$condition)) , nchar(as.character(study1$condition)) ) #trims unneeded text
study1$condition_code = as.numeric(study1$condition)  #useful for attention check coding
study1$condition = as.factor(ifelse(study1$condition==1, "prosocial", ifelse(study1$condition==2, "antisocial", "neutral")))
study1$prosocial_condition = ifelse(study1$condition=="prosocial", 1, 0) 
```

#### Study 2
Study 2 included two video conditions: 1-pro, 2-ant
```{r}
study2$condition = as.factor(ifelse(study2$condition=="pro", "prosocial", ifelse(study2$condition=="ant", "antisocial", "ERROR")))
study2$prosocial_condition = ifelse(study2$condition=="prosocial", 1, 0) 
study2$condition_code = ifelse(study2$prosocial_condition==1, 1, 2) #useful for attention check coding
```

#### Study 3
Study 3 included prosocial, antisocial, and neutral conditions.
```{r}
study3$prosocial_condition = ifelse(study3$condition=="prosocial", 1, 0) 
```

####Study 4
Study 4 was a 2x2 design with cells:  
`r levels(study4$branch)`
```{r}
# this could be cleaner, but it works
study4$condition=as.factor(strtrim(study4$branch, 6))
levels(study4$condition)
study4$prosocial_condition = ifelse(study4$condition=="Elevat", 1, ifelse(study4$condition=="Contro", 0, NA))

study4$optin=as.numeric(grepl("Opt-in", study4$branch))
```

####Study 5
Study 5 just has Parkour (control) and Thai (prosocial_condition) conditions.  

```{r}
study5$prosocial_condition = ifelse(study5$condition=="Thai", 1, ifelse(study5$condition=="Parkour", 0, NA))
```

####Study 6 & 7 
Study 6 and 7 were a 2 (order) x 2 (condition) design with cells: `r levels(study6$Cell)` and `r levels(study7$Cell)`. Separate these out, in text and dummy coding forms:
```{r}
study6$condition = substring(study6$Cell, 3)
study6$prosocial_condition = ifelse(study6$condition=="thai", 1, ifelse(study6$condition=="parkour", 0, NA))

study6$order = substring(study6$Cell, 1, 1)
study6$orderA =  ifelse(study6$order=="A", 1, ifelse(study6$order=="B", 0, NA))

study7$condition = substring(study7$Cell, 7)
study7$prosocial_condition = ifelse(study7$condition=="thai", 1, ifelse(study7$condition=="parkour", 0, NA))

study7$order = substring(study7$Cell, 1, 5)
study7$orderA =  ifelse(study7$order=="A1 B2", 1, ifelse(study7$order=="B1 A2", 0, NA))
```

### Condensing variables
Sometimes my SurveyGizmo set up requires the same question to be reported in more than one column, typically because of branching logic. In this section I condense those into a single variable.

*I bet some package has a better way to do this.*

####Study 4
The "opt-in" vs. "opt-out" framings presented basically the same charitable donation question.

```{r}
study4$charity=rowMeans(study4[,c("optin","optout")],na.rm=TRUE) # this is goofy but it works
```

#### Study 6
Identical tems were presented before and/or after the manipulation.

For example, the variables "a.honest.pre" and "a.honest.post" refer to the same attitude scale item (hence "a.xxxxxxx") and whether it was asked pre- or post-manipulation. It was asked before the manipulation in order A, after in order B. I just want one variable "a.honest" and use variable "order" to separate pre-/post-manipulation when needed. The code below condenses all relevant variables.

*I'm not sure that this is the best way to proceed, from the perspective of data structure, but I'm going ahead for now. I'm probably choosing brute force over efficiency, but maybe the efficient solution will emerge.*

*More importantly, there's probably some text-function based way (package "stringr" ?) to simplify the shit out of this--this monstrosity is another instance of using brute force over efficiency for now.*

*I bet some package makes this simpler. {tidyR} maybe?*

```{r}
#Attitude: In block A, before the video. After in block B. (attitude.preA.postB)
study6$a.honest = ifelse(study6$order=="A", study6$a.honest.pre, study6$a.honest.post)
study6$a.malicious = ifelse(study6$order=="A", study6$a.malicious.pre, study6$a.malicious.post)
study6$a.noadvantage = ifelse(study6$order=="A", study6$a.noadvantage.pre, study6$a.noadvantage.post)
study6$a.pessimistic = ifelse(study6$order=="A", study6$a.pessimistic.pre, study6$a.pessimistic.post)
study6$a.befair = ifelse(study6$order=="A", study6$a.befair.pre, study6$a.befair.post)
study6$a.nothelpful = ifelse(study6$order=="A", study6$a.nothelpful.pre, study6$a.nothelpful.post)
study6$a.trustworthy = ifelse(study6$order=="A", study6$a.trustworthy.pre, study6$a.trustworthy.post)
study6$a.notgood  = ifelse(study6$order=="A", study6$a.notgood.pre,study6$a.notgood.post)
study6$a.dontbecareful  = ifelse(study6$order=="A", study6$a.dontbecareful.pre,study6$a.dontbecareful.post)
study6$a.inottrust  = ifelse(study6$order=="A", study6$a.inottrust.pre,study6$a.inottrust.post)
study6$a.lookoutothers = ifelse(study6$order=="A", study6$a.lookoutothers.pre,study6$a.lookoutothers.post)
study6$a.peoplenottrust = ifelse(study6$order=="A", study6$a.peoplenottrust.pre,study6$a.peoplenottrust.post)
study6$a.good = ifelse(study6$order=="A", study6$a.good.pre,study6$a.good.post)
study6$a.careonly = ifelse(study6$order=="A", study6$a.careonly.pre,study6$a.careonly.post)
study6$a.awful = ifelse(study6$order=="A", study6$a.awful.pre, study6$a.awful.post)
#attitude.preB.postA
study6$a.dishonest = ifelse(study6$order=="B", study6$a.dishonest.pre,study6$a.dishonest.post)  
study6$a.kind = ifelse(study6$order=="B", study6$a.kind.pre,study6$a.kind.post)  
study6$a.takeadvantage = ifelse(study6$order=="B", study6$a.takeadvantage.pre,study6$a.takeadvantage.post)  
study6$a.optimistic = ifelse(study6$order=="B", study6$a.optimistic.pre,study6$a.optimistic.post)  
study6$a.notfair = ifelse(study6$order=="B", study6$a.notfair.pre,study6$a.notfair.post)  
study6$a.helpful = ifelse(study6$order=="B", study6$a.helpful.pre,study6$a.helpful.post)  
study6$a.nottrustworthy = ifelse(study6$order=="B", study6$a.nottrustworthy.pre, study6$a.nottrustworthy.post)
study6$a.canbegood = ifelse(study6$order=="B", study6$a.canbegood.pre,study6$a.canbegood.post)  
study6$a.becareful = ifelse(study6$order=="B", study6$a.becareful.pre,study6$a.becareful.post)  
study6$a.itrust = ifelse(study6$order=="B", study6$a.itrust.pre,study6$a.itrust.post)
study6$a.lookoutself = ifelse(study6$order=="B", study6$a.lookoutself.pre, study6$a.lookoutself.post)  
study6$a.peopletrust = ifelse(study6$order=="B", study6$a.peopletrust.pre, study6$a.peopletrust.post) 
study6$a.evil = ifelse(study6$order=="B", study6$a.evil.pre,study6$a.evil.post)  
study6$a.admirable = ifelse(study6$order=="B", study6$a.admirable.pre,study6$a.admirable.post)  
study6$a.caremore = ifelse(study6$order=="B", study6$a.caremore.pre,study6$a.caremore.post)  

#prosocial.motives.preA.postB = c("e.wanttohelp","e.feelclose","e.newhelp")
study6$e.wanttohelp = ifelse(study6$order=="A", study6$e.wanttohelp.pre, study6$e.wanttohelp.post)
study6$e.feelclose = ifelse(study6$order=="A", study6$e.feelclose.pre, study6$e.feelclose.post)
study6$e.newhelp = ifelse(study6$order=="A", study6$e.newhelp.pre, study6$e.newhelp.post)

#prosocial.motives.preB.postA = c("e.donate","e.worlddeserves","e.betterperson")
study6$e.donate = ifelse(study6$order=="B", study6$e.donate.pre, study6$e.donate.post)
study6$e.worlddeserves = ifelse(study6$order=="B", study6$e.worlddeserves.pre, study6$e.worlddeserves.post)
study6$e.betterperson = ifelse(study6$order=="B", study6$e.betterperson.pre, study6$e.betterperson.post)

#positive.affect = c("pa.happy", "pa.amused","pa.delighted","pa.entertained")

#positive.affect.preA.postB = c("pa.happy", "pa.amused") 
study6$pa.happy = ifelse(study6$order=="A", study6$pa.happy.pre, study6$pa.happy.post)
study6$pa.amused = ifelse(study6$order=="A", study6$pa.amused.pre, study6$pa.amused.post)

#positive.affect.preB.postA = c("pa.delighted","pa.entertained") 
study6$pa.delighted = ifelse(study6$order=="B", study6$pa.delighted.pre, study6$pa.delighted.post)
study6$pa.entertained = ifelse(study6$order=="B",study6$pa.entertained.pre, study6$pa.entertained.post)

#elevating.emotions.pre  = c("e.uplifted.pre","e.compassion.pre","e.admiration.pre","e.heartwarming.pre","e.moved.pre","e.touched.pre","e.inspired.pre")
study6$e.uplifted.pre =ifelse(study6$order=="A", study6$e.uplifted.pre1 ,study6$e.uplifted.pre2)
study6$e.compassion.pre =ifelse(study6$order=="A", study6$e.compassion.pre1 ,study6$e.compassion.pre2)
study6$e.admiration.pre =ifelse(study6$order=="A", study6$e.admiration.pre1 ,study6$e.admiration.pre2)
study6$e.heartwarming.pre=ifelse(study6$order=="A", study6$e.heartwarming.pre1 ,study6$e.heartwarming.pre2)
study6$e.moved.pre =ifelse(study6$order=="A", study6$e.moved.pre1 ,study6$e.moved.pre2)
study6$e.touched.pre =ifelse(study6$order=="A", study6$e.touched.pre1 ,study6$e.touched.pre2)
study6$e.inspired.pre =ifelse(study6$order=="A", study6$e.inspired.pre1 ,study6$e.inspired.pre2)

#elevating.emotions.post = c("e.uplifted.post","e.compassion.post","e.admiration.post","e.heartwarming.post","e.moved.post","e.touched.post","e.inspired.post")
study6$e.uplifted.post =ifelse(study6$order=="A", study6$e.uplifted.post1 ,study6$e.uplifted.post2)
study6$e.compassion.post =ifelse(study6$order=="A", study6$e.compassion.post1 ,study6$e.compassion.post2)
study6$e.admiration.post =ifelse(study6$order=="A", study6$e.admiration.post1 ,study6$e.admiration.post2)
study6$e.heartwarming.post=ifelse(study6$order=="A", study6$e.heartwarming.post1 ,study6$e.heartwarming.post2)
study6$e.moved.post =ifelse(study6$order=="A", study6$e.moved.post1 ,study6$e.moved.post2)
study6$e.touched.post =ifelse(study6$order=="A", study6$e.touched.post1 ,study6$e.touched.post2)
study6$e.inspired.post =ifelse(study6$order=="A", study6$e.inspired.post1 ,study6$e.inspired.post2)
```

####Study 7

```{r}
study7$a.honest = ifelse(study7$orderA=="1", study7$a.honest.pre, study7$a.honest.post)
study7$a.malicious = ifelse(study7$orderA=="1", study7$a.malicious.pre, study7$a.malicious.post)
study7$a.noadvantage = ifelse(study7$orderA=="1", study7$a.noadvantage.pre, study7$a.noadvantage.post)
study7$a.pessimistic = ifelse(study7$orderA=="1", study7$a.pessimistic.pre, study7$a.pessimistic.post)
study7$a.befair = ifelse(study7$orderA=="1", study7$a.befair.pre, study7$a.befair.post)
study7$a.nothelpful = ifelse(study7$orderA=="1", study7$a.nothelpful.pre, study7$a.nothelpful.post)
study7$a.trustworthy = ifelse(study7$orderA=="1", study7$a.trustworthy.pre, study7$a.trustworthy.post)
study7$a.notgood  = ifelse(study7$orderA=="1", study7$a.notgood.pre,study7$a.notgood.post)
study7$a.dontbecareful  = ifelse(study7$orderA=="1", study7$a.dontbecareful.pre,study7$a.dontbecareful.post)
study7$a.inottrust  = ifelse(study7$orderA=="1", study7$a.inottrust.pre,study7$a.inottrust.post)
study7$a.lookoutothers = ifelse(study7$orderA=="1", study7$a.lookoutothers.pre,study7$a.lookoutothers.post)
study7$a.peoplenottrust = ifelse(study7$orderA=="1", study7$a.peoplenottrust.pre,study7$a.peoplenottrust.post)
study7$a.good = ifelse(study7$orderA=="1", study7$a.good.pre,study7$a.good.post)
study7$a.notadmirable = ifelse(study7$orderA=="1", study7$a.notadmirable.pre, study7$a.notadmirable.post)
study7$a.careonly = ifelse(study7$orderA=="1", study7$a.careonly.pre,study7$a.careonly.post)

#attitude.preB.postA
study7$a.dishonest = ifelse(study7$orderA=="0", study7$a.dishonest.pre,study7$a.dishonest.post)  
study7$a.kind = ifelse(study7$orderA=="0", study7$a.kind.pre, study7$a.kind.post)
study7$a.takeadvantage = ifelse(study7$orderA=="0", study7$a.takeadvantage.pre,study7$a.takeadvantage.post)  
study7$a.optimistic = ifelse(study7$orderA=="0", study7$a.optimistic.pre,study7$a.optimistic.post)  
study7$a.notfair = ifelse(study7$orderA=="0", study7$a.notfair.pre,study7$a.notfair.post)  
study7$a.helpful = ifelse(study7$orderA=="0", study7$a.helpful.pre,study7$a.helpful.post)  
study7$a.nottrustworthy = ifelse(study7$orderA=="0", study7$a.nottrustworthy.pre, study7$a.nottrustworthy.post)
study7$a.canbegood = ifelse(study7$orderA=="0", study7$a.canbegood.pre,study7$a.canbegood.post)  
study7$a.becareful = ifelse(study7$orderA=="0", study7$a.becareful.pre,study7$a.becareful.post)  
study7$a.itrust = ifelse(study7$orderA=="0", study7$a.itrust.pre,study7$a.itrust.post)
study7$a.lookoutself = ifelse(study7$orderA=="0", study7$a.lookoutself.pre, study7$a.lookoutself.post)  
study7$a.peopletrust = ifelse(study7$orderA=="0", study7$a.peopletrust.pre, study7$a.peopletrust.post) 
study7$a.littlegood = ifelse(study7$orderA=="0", study7$a.littlegood.pre,study7$a.littlegood.post)  
study7$a.admirable = ifelse(study7$orderA=="0", study7$a.admirable.pre,study7$a.admirable.post)  
study7$a.caremore = ifelse(study7$orderA=="0", study7$a.caremore.pre,study7$a.caremore.post)  
```

### Reverse-scoring
Many analyses are more convenient if elevation/idealism/prosociality is higher.

####Study 2
```{r}
study2$shape2 <-  6-study2$shape2
```

####Study3
Here I'm reversing attitude items such that idealism is always higher. Not bothering with the many other items.
```{r}
names(study3)[5:14] = c(
  "a.dishonest",
  "a.kind",
  "a.itrust",
  "a.nottrust",
  "a.trustworthy",
  "a.takeadvantage",
  "a.befair",
  "a.careful",
  "a.helpful",
  "a.lookingout"
  )
study3$a.dishonest<-8-study3$a.dishonest 
study3$a.nottrust<-8-study3$a.nottrust
study3$a.takeadvantage<-8-study3$a.takeadvantage
study3$a.careful<-8-study3$a.careful
study3$a.lookingout<-8-study3$a.lookingout
```

####Study 4
```{r}
study4$a.dishonest=8-study4$a.dishonest 
study4$a.nottrust=8-study4$a.nottrust
study4$a.takeadvantage=8-study4$a.takeadvantage
study4$a.careful=8-study4$a.careful
study4$a.lookingout=8-study4$a.lookingout
study4$shape2=6-study4$shape2
```

####Study 5
```{r}
study5$a.dishonest=8-study5$a.dishonest 
study5$a.nottrust=8-study5$a.nottrust
study5$a.takeadvantage=8-study5$a.takeadvantage
study5$a.careful=8-study5$a.careful
study5$a.lookingout=8-study5$a.lookingout
study5$shape=6-study5$shape
```

####Study 6
```{r}
study6$a.malicious = 8 - study6$a.malicious
study6$a.pessimistic = 8 - study6$a.pessimistic
study6$a.nothelpful = 8 - study6$a.nothelpful
study6$a.notgood = 8 - study6$a.notgood
study6$a.inottrust = 8 - study6$a.inottrust
study6$a.peoplenottrust = 8 - study6$a.peoplenottrust
study6$a.dishonest = 8 - study6$a.dishonest
study6$a.takeadvantage = 8 - study6$a.takeadvantage
study6$a.notfair = 8 - study6$a.notfair
study6$a.nottrustworthy = 8 - study6$a.nottrustworthy
study6$a.becareful = 8 - study6$a.becareful
study6$a.lookoutself = 8 - study6$a.lookoutself

study6$a.evil = 3 - study6$a.evil
study6$a.careonly = 3 - study6$a.careonly
study6$a.awful = 3 - study6$a.awful

study6$shape = 6 - study6$shape
```

####Study 7
```{r}
study7$a.malicious = 8 - study7$a.malicious
study7$a.pessimistic = 8 - study7$a.pessimistic
study7$a.nothelpful = 8 - study7$a.nothelpful
study7$a.notgood = 8 - study7$a.notgood
study7$a.inottrust = 8 - study7$a.inottrust
study7$a.peoplenottrust = 8 - study7$a.peoplenottrust
study7$a.notadmirable = 8 - study7$a.notadmirable
study7$a.careonly = 8 - study7$a.careonly

study7$a.dishonest = 8 - study7$a.dishonest
study7$a.takeadvantage = 8 - study7$a.takeadvantage
study7$a.notfair = 8 - study7$a.notfair
study7$a.nottrustworthy = 8 - study7$a.nottrustworthy
study7$a.becareful = 8 - study7$a.becareful
study7$a.lookoutself = 8 - study7$a.lookoutself
study7$a.littlegood = 8 - study7$a.littlegood

study7$shape1 = 6 - study7$shape1
study7$shape2 = 6 - study7$shape2
```



### Response options clarifications
I don't see a SurveyMonkey option to export the survey including the response options, so I had to manually log in and look up what the numeric reporting values are. 

```{r}
#after study 1 we dropped the words and made the others likert-like rather than binary
study1$word1 = as.factor(ifelse(study1$word1 == 1, "kiki", "bouba"))
study1$word2 = as.factor(ifelse(study1$word2 == 1, "maluma", "taketa"))
study1$shape1 = as.factor(ifelse (study1$shape1 == 1, "spike explosion", "cloud"))
study1$shape2 = as.factor(ifelse (study1$shape2 == 1, "boomerang", "spikerang"))
study1$circles = as.factor(ifelse (study1$circles == 1, "close", "far"))

#these are basicallly binary endorsements
study1$circle.reason1 = as.numeric(is.na(study1$circle.reason1) == FALSE)
study1$circle.reason2 = as.numeric(is.na(study1$circle.reason2) == FALSE)
study1$circle.reason3 = as.numeric(is.na(study1$circle.reason3) == FALSE)
study1$circle.reason4 = as.numeric(is.na(study1$circle.reason4) == FALSE)
study1$circle.reason5 = as.numeric(is.na(study1$circle.reason5) == FALSE)
study1$circle.reason6 = as.numeric(is.na(study1$circle.reason6) == FALSE)

study2$circle.reason1 = as.numeric(is.na(study2$circle.reason1) == FALSE)
study2$circle.reason2 = as.numeric(is.na(study2$circle.reason2) == FALSE)
study2$circle.reason3 = as.numeric(is.na(study2$circle.reason3) == FALSE)
study2$circle.reason4 = as.numeric(is.na(study2$circle.reason4) == FALSE)
study2$circle.reason5 = as.numeric(is.na(study2$circle.reason5) == FALSE)
study2$circle.reason6 = as.numeric(is.na(study2$circle.reason6) == FALSE)
```

### Demographic variable clean-up.
####Height
```{r}
study1$halfinch[is.na(study1$halfinch)]=0
study2$halfinch[is.na(study2$halfinch)]=0
study3$halfinch[is.na(study3$halfinch)]=0

study1$height= (12*(study1$feet+2)) + study1$inch-1 + ifelse(study1$halfinch ==2, 0.5, 0) #1st option for # of feet was 3, for inches 0
study2$height= (12*(study2$feet+2)) + study2$inch-1 + ifelse(study2$halfinch ==2, 0.5, 0) #1st option for # of feet was 3, for inches 0
study3$height= (12*(study3$feet+2)) + study3$inch-1 + ifelse(study3$halfinch ==2, 0.5, 0) #1st option for # of feet was 3, for inches 0
study4$height= (12*study4$feet) + study4$inches
study5$height= (12*study5$feet) + study5$inches
study6$height= (12*study6$feet) + study6$inches
study7$height= (12*study7$feet) + study7$inches

study1$height[study1$height==45]=NA #extreme value
study6$height[study6$height==48]=NA #extreme value

study1$height_z = scale(study1$height, center=T, scale=T)
study2$height_z = scale(study2$height, center=T, scale=T)
study3$height_z = scale(study3$height, center=T, scale=T)
study4$height_z = scale(study4$height, center=T, scale=T)
study5$height_z = scale(study5$height, center=T, scale=T)
study6$height_z = scale(study6$height, center=T, scale=T)
study7$height_z = scale(study7$height, center=T, scale=T)
```

#### Age
```{r}
study1$age[study1$age==1992]= 2015-1992 #i assume they entered birth year
study1$age[study1$age==1994]= 2015-1994 #i assume they entered birth year
study2$age[study2$age==256]=NA #extreme value
study4$age[study4$age==300]=NA #extreme value
study5$age[study5$age==2]=NA #extreme value

study1$age_z = scale(study1$age, center=T, scale=T)
study2$age_z = scale(study2$age, center=T, scale=T)
study3$age_z = scale(study3$age, center=T, scale=T)
study4$age_z = scale(study4$age, center=T, scale=T)
study5$age_z = scale(study5$age, center=T, scale=T)
study6$age_z = scale(study6$age, center=T, scale=T)
study7$age_z = scale(study7$age, center=T, scale=T)
```

####Political orientation
```{r}
study4$political[study4$political=="na" | study4$political==""]=NA
study5$political[study5$political=="na" | study5$political==""]=NA
study6$political[study6$political=="na" | study6$political==""]=NA 
study7$political[study7$political=="na" | study7$political==""]=NA 

study4$political=droplevels(study4$political)
study4$political =as.numeric(study4$political)

study5$political=droplevels(study5$political)
study5$political =as.numeric(study5$political)

study6$political=droplevels(study6$political)
study6$political =as.numeric(study6$political)

study7$political=droplevels(study7$political)
study7$political =as.numeric(study7$political)
```

####Sex
Participants have the option to skip this item or pick "other / prefer not to say." Vast majority pick male or female, so those are the only analyzable categories.
```{r}
study1$sex = as.factor(study1$sex)
study1$sex[study1$sex==0] = NA 
study1$sex=droplevels(study1$sex) 

study2$sex = as.factor(study2$sex)

study4$sex[study4$sex==""|study4$sex=="na"]=NA
study4$sex=droplevels(study4$sex) 

study5$sex[study5$sex==""|study5$sex=="na"]=NA
study5$sex=droplevels(study5$sex) 

study6$sex[study6$sex==""|study6$sex=="na"]=NA
study6$sex=droplevels(study6$sex) 

study7$sex[study7$sex==""|study7$sex=="na"]=NA
study7$sex=droplevels(study7$sex) 
```

#### Income
I could probably adjust the input for this item to simplify this cleaning operation.
```{r}
#playground
#table(study4$income, useNA="always")
#a = as.character(study4$income)
#b=as.numeric(gsub(",", "", a))
#a[is.na(b)]
#a

study1$income[study1$income==""] = NA

study4$income[study4$income=="rather not"|study4$income==""]=NA
study4$income[study4$income=="$20,000 "] = 20000
study5$income[study5$income=="no clue"|study5$income==""]=NA
study6$income[study6$income=="homeless graduate student"|study6$income==""]=NA
study7$income[study7$income==""]=NA

#study1$income = droplevels(study1$income) #not sure if i need this -- test next time i run in order
study4$income = droplevels(study4$income)
study5$income = droplevels(study5$income)
study6$income = droplevels(study6$income)
study7$income = droplevels(study7$income)

study1$income=as.numeric(gsub(",", "", study1$income))
study4$income=as.numeric(gsub(",", "", study4$income))
study5$income=as.numeric(gsub(",", "", study5$income))
study6$income=as.numeric(gsub(",", "", study6$income))
study7$income=gsub(",", "", study7$income)
study7$income=as.numeric(gsub("\\$", "", study7$income))

#used as demographic control variables, so scale them to make effects interpretable
study1$income_z=as.numeric(scale(study1$income,scale=TRUE,center=TRUE))
study2$income_z=as.numeric(scale(study2$income,scale=TRUE,center=TRUE))
study3$income_z=as.numeric(scale(study3$income,scale=TRUE,center=TRUE))
study4$income_z=as.numeric(scale(study4$income,scale=TRUE,center=TRUE))
study5$income_z=as.numeric(scale(study5$income,scale=TRUE,center=TRUE))
study6$income_z=as.numeric(scale(study6$income,scale=TRUE,center=TRUE))
study7$income_z=as.numeric(scale(study7$income,scale=TRUE,center=TRUE))
```

####Education
In each case I condense response codes 1 and 2 into "high school or less" and 6, 7, and 8 into "some grad school or more."
```{r}
study1$hs_less = as.numeric(study1$education < 3)
study1$some_college = as.numeric(study1$education == "3")
study1$AA = as.numeric(study1$education == "4")
study1$BA = as.numeric(study1$education == "5")
study1$grad_school = as.numeric(study1$education > 5)

study2$hs_less = as.numeric(study2$education < 3)
study2$some_college = as.numeric(study2$education == "3")
study2$AA = as.numeric(study2$education == "4")
study2$BA = as.numeric(study2$education == "5")
study2$grad_school = as.numeric(study2$education > 5)

study3$hs_less = as.numeric(study3$education < 3)
study3$some_college = as.numeric(study3$education == "3")
study3$AA = as.numeric(study3$education == "4")
study3$BA = as.numeric(study3$education == "5")
study3$grad_school = as.numeric(study3$education > 5)

study4$hs_less = as.numeric(study4$education < 3)
study4$some_college = as.numeric(study4$education == "3")
study4$AA = as.numeric(study4$education == "4")
study4$BA = as.numeric(study4$education == "5")
study4$grad_school = as.numeric(study4$education > 5)

study5$hs_less = as.numeric(study5$education < 3)
study5$some_college = as.numeric(study5$education == "3")
study5$AA = as.numeric(study5$education == "4")
study5$BA = as.numeric(study5$education == "5")
study5$grad_school = as.numeric(study5$education > 5)

study6$hs_less = as.numeric(study6$education < 3)
study6$some_college = as.numeric(study6$education == "3")
study6$AA = as.numeric(study6$education == "4")
study6$BA = as.numeric(study6$education == "5")
study6$grad_school = as.numeric(study6$education > 5)

study7$hs_less = as.numeric(study7$education < 3)
study7$some_college = as.numeric(study7$education == "3")
study7$AA = as.numeric(study7$education == "4")
study7$BA = as.numeric(study7$education == "5")
study7$grad_school = as.numeric(study7$education > 5)

```
#### Ethnicity
Majority of each sample is white, so for now just code for white / non-white.
```{r}
study1$nonwhite = as.numeric(study1$ethnic!=1)
study2$nonwhite = as.numeric(study2$ethnic!=1)
study4$nonwhite = as.numeric(study4$ethnic!="White")
study5$nonwhite = as.numeric(study5$ethnic!="White")
study6$nonwhite = as.numeric(study6$ethnic!="White")
study7$nonwhite = as.numeric(study7$ethnic!="White")
```
#### Pregnancy
The reported number of pregnancies was small, so this variable is ignored.

## Exclusions

### Examine information relevant to potential exclusions
#### Duplicate IP
Study 1 & 2 have some of these. But we used the unique Turker method so repeats must have used different MTurk identities.

What were settings on other studies?

#### Finishing time
The threshold for minimum acceptible time is somewhat arbitrary. In most conditions (expection: Study 4 Control) participants had to watch a 3 minute video. 

```{r}
library(lubridate)

study1$time=as.numeric(difftime(mdy_hms(study1$EndDate), mdy_hms(study1$StartDate)))/60
study2$time=as.numeric(difftime(mdy_hms(study2$EndDate), mdy_hms(study2$StartDate)))/60
study3$time=as.numeric(difftime(mdy_hm(study3$EndDate), mdy_hm(study3$StartDate)))
study4$time=as.numeric(difftime(mdy_hm(study4$Date.Submitted), mdy_hm(study4$Time.Started)))
study5$time=as.numeric(difftime(mdy_hm(study5$Date.Submitted), mdy_hm(study5$Time.Started)))
study6$time=as.numeric(difftime(mdy_hm(study6$Date.Submitted), mdy_hm(study6$Time.Started)))
study7$time=as.numeric(difftime(ymd_hms(study7$Date.Submitted), ymd_hms(study7$Time.Started))) 
#I have just noticed that the study7 date/time format is the source format for all studies, but using Excel in my workflow eliminated the seconds data. I used LibreCalc for study7 and study1 and retain the seconds. Probably not worth recoding old data for a slightly more precise time estimate, but good to know for future.

hist(study1$time, breaks=300)
hist(study2$time, breaks=300)
table(study3$time)
table(study4$time, by=study4$condition)
table(study5$time)
table(study6$time)
hist(study7$time, breaks=500)
``` 

The distributions of responses above suggest that studies 1, 2, 4 and 5 would require at minimum another 2 minutes to finish all questions, Studies 4 and 6 would require 3 more minutes, Study 7 requires 3.5 more. These seem intuitively reasonable, so later I will implement those cutoffs.

#### Missing items
This is important for study 1.

Later studies handle this in relevant analyses.

```{r}
# study1
study1$missing = rowSums(is.na(study1)) #keep missing < 18
study2$missing = rowSums(is.na(study2)) #keep missing < 18

#later studies?
#count up missing items from the key scales.
# study6$missing = rowSums(is.na(study6[,c(attitude, elevating.emotions.pre, elevating.emotions.post, positive.affect, prosocial.motives)]))
```

#### Playback problems
##### Study 4
Exclusions based on playback problems were done manually in spreadsheet software.

##### Study 5
`r length(subset(study5$playback, study5$playback!=1))` people reported playback problems. Of these, `r length(subset(study5$playback, study5$playback==2))` were reported as "minor." Here are the descriptions they provided:
```{r}
as.character(subset(study5, study5$playback==2)$playback.problems)
```
Based on those descriptions, I judge that none of these cases need to be eliminated. "Major problems" were reported by `r length(subset(study5$playback, study5$playback==3))` participants. They described them:
```{r}
as.character(subset(study5, study5$playback==3)$playback.problems)
```
Based on these descriptions, I judge that all these cases should be eliminated.

##### Study 6
`r length(subset(study6$playback, study6$playback!=1))` people reported playback problems. Of these, `r length(subset(study6$playback, study6$playback==2))` were reported as "minor." Here are the descriptions they provided:
```{r}
as.character(subset(study6, study6$playback==2)$playback.problems)
```
Based on those descriptions, I judge that none of these cases need to be eliminated. "Major problems" were reported by `r length(subset(study6$playback, study6$playback==3))` participants. They described them:
```{r}
as.character(subset(study6, study6$playback==3)$playback.problems)
```
Based on these descriptions, I judge that all these cases should be eliminated.


```{r include = FALSE}
# Picking out those playback problem subsets resulted in some weird NA-related errors. The as.character() move superficially solved them. And they look accurate too. But there's some glitch here I don't understand. I tried the fixes below without success.
#study6$Response.ID = as.character(study6$Response.ID)
#study6$Response.ID = droplevels(study6$Response.ID)
```


```{r include=FALSE}
#There might be slight trends towards less missing data and less quick finishing in thai condition.
table(study6$missing[study6$condition=="parkour"],useNA = "always")
table(study6$missing[study6$condition=="thai"],useNA = "always")

mean(subset(study6$missing, study6$condition == "thai")) 
mean(subset(study6$missing, study6$condition == "parkour")) 

table(study6$time[study6$condition=="parkour"], useNA = "always")
table(study6$time[study6$condition=="thai"], useNA = "always")
```

##### Study 7
`r length(subset(study7$playback, study7$playback!=1))` people reported playback problems. Of these, `r length(subset(study7$playback, study7$playback==2))` were reported as "minor." Here are the descriptions they provided:
```{r}
as.character(subset(study7, study7$playback==2)$playback.problems)
```
Based on those descriptions, I judge that one should be reclassified as a major problem. 
```{r}
study7[study7$playback.problems=="Second video didn't really place but I still enjoyed the blank screen made me feel all warm and good inside.",]$playback = 3
```



"Major problems" were reported by `r length(subset(study7$playback, study7$playback==3))` participants. They described them:
```{r}
as.character(subset(study7, study7$playback==3)$playback.problems)
```
Based on these descriptions, I judge that all these cases should be eliminated.






### Exclude in data set

This code makes all of the eliminations.

```{r}
study1 = subset (study1, study1$check1==1 & study1$check2==26 & study1$check3==1+study1$condition_code & study1$missing < 18)
study2 = subset (study2, study2$check_alphabet==26 & study2$check_video==1+study2$condition_code & study2$missing < 18)
study3 = subset (study3, (study3$vidcheck==2 & study3$condition=="prosocial") | (study3$vidcheck==3 & study3$condition== "antisocial") | (study3$vidcheck==4 & study3$condition=="neutral")) #i did this in excel but this could replace that

study4 = subset (study4, study4$check == "blue")
study5 = subset (study5, study5$check == "blue")
study6 = subset (study6, study6$check == "blue")
study7 = subset (study7, study7$check == "blue")

study4 = subset(study4, study4$device != "Phone")
study5 = subset(study5, study5$device != "Phone")
study6 = subset(study6, study6$device != "Phone")
study7 = subset(study7, study7$device != "Phone")

study1=subset(study1, study1$time>5)
study2=subset(study2, study2$time>5)
study3=subset(study3, study3$time>6)
study4=subset(study4, study4$time>1 & study4$condition=="Contro" | study4$time>4 & study4$condition=="Elevat")
study5=subset(study5, study5$time>4)
study6=subset(study6, study6$time>5) 
study7=subset(study7, study7$time>6.5)

# drops "serious playback problems"" from the data set.
study4=subset(study4, study4$playback==1 | study4$playback==2 | is.na(study4$playback)=="TRUE")
study5=subset(study5, study5$playback==1 | study5$playback==2 | is.na(study5$playback)=="TRUE")
study6=subset(study6, study6$playback==1 | study6$playback==2 | is.na(study6$playback)=="TRUE")
study7=subset(study7, study7$playback==1 | study7$playback==2 | is.na(study7$playback)=="TRUE")

#drop unneeded variables
study1 = subset(study1, select= -c(CollectorID, currency, complete))
```

Final Samples

```{r}

final_sample = list(
  "Study 1" = dim(study1)[1],
  "Study 2" = dim(study2)[1],
  "Study 3" = dim(study3)[1],
  "Study 4" = dim(study4)[1],
  "Study 5" = dim(study5)[1],
  "Study 6" = dim(study6)[1],
  "Study 7" = dim(study7)[1]
  )

starting_sample
final_sample
```


(The sum of the exclusion types in each category may not equal the difference between starting and ending sample because some participants registered multiple disqualifiers.) Here are revelant condition breakdowns

```{r}
summary(study1$condition)
summary(study2$condition)
summary(study3$condition)
summary(study4$branch)
summary(study5$condition)
summary(study6$Cell)
summary(study7$Cell)
```

## Write clean files for analysis
```{r}
write.csv(study1, "study1_cleandata.csv")
write.csv(study2, "study2_cleandata.csv")
write.csv(study3, "study3_cleandata.csv")
write.csv(study4, "study4_cleandata.csv")
write.csv(study5, "study5_cleandata.csv")
write.csv(study6, "study6_cleandata.csv")
write.csv(study7, "study7_cleandata.csv")
```

## Ways to improve this script
1. For some of the studies, I did some cleaning in excel. Do all of that here.
2. Variable short/long names files for every study.
3. These final clean files aren't "tidy" (in the Wickham sense). Package "tidyR" would help.
4. Some variable names still include "." (the character for period) -- shift to using "_" (underscore) to avoid confusion with other programming languages where periods are special
5. Is there some better way to do variable names overall?
6. The code to drop cases is separate from the code that informs those decisions. They should probably be closer together UNLESS we care how many cases meet each exclusion criteria (because if we eliminate based on criteria 1 before examining criteria 2, our count of violations of 2 would omit cases already eliminated based on criteria 1.)
7. Package "dplyr" would likely help improve efficency.
